### 2023-03-02 manuscript.Rmd
last changes

Based on four principles of a more responsible research assessment in academic hiring and promotion processes [@schonbrodt2022responsible], @gartner2022responsible suggested a concrete evaluation scheme for published manuscripts, reusable datasets, and research software.
We strongly support the increased emphasis on research software as a creditable and commendable scientific contribution.
Why are research software contributions important scientific contributions?
We would like to respond with a quote from the Science Code Manifesto [@manifesto]:
"Software is a cornerstone of science. Without software, twenty-first century science would be impossible."
and
"Software is an essential research product, and the effort to produce, maintain, adapt, and curate code must be recognized."
However, despite the heavy reliance on computational infrastructure, the current academic infrastructure does not adequately incentivise software development and, specifically, good software engineering practice [@baxter2012research].
In line with principle 3 of @schonbrodt2022responsible, we suggest that criteria for research software contributions must capture two major dimensions: rigour and impact.
Impact measures whether the scholarly effort, implementation, and dissemination actually had a visible effect on the field.
rigour means implementing high standards and best practices for ensuring transparency, correctness, and reusability of a piece of software.
By setting a high bar of rigour in research software assessment in academic hiring and promotion, we hope to foster the creation of better software and, thus, better science.
From this perspective, we comment on some indicators of the proposed evaluation scheme for research software contributions (Table 3 of @gartner2022responsible).

@@

## ID 5: Date of first fully functional public version/ID 9: Citations

@@

Both ID 5 and 6 are difficult to ascertain because it is not always clear when updates are considered 'substantive' or software 'fully functional'.
We appreciate the importance of assessing maintenance as part of rigour.
To assess this, we propose to provide a simple checkbox, in which the author indicates whether a scientific software package is actively maintained (e.g., the software has a regular release cycle or an update within the last six months).
In addition, authors have a chance to explain why their active maintenance may be different from these guidelines.

@@

At present, whether a piece of software is open source is not evaluated in the prescreening phase.
However, an open license is central for assessing both rigour and (potential) impact, and should be part of the phase I assessment.
Above, we already discussed the different ways research software can have an impact --- not only by direct usage, but also by reusing software in other packages.
Open-source software makes broader impact more likely and is a prerequisite for full transparency and reproducibility [@peikert2021reproducible].
In addition, many aspects of rigour are impossible to evaluate for closed-source software --- for example, whether it is well-tested or bugs have been fixed.
Therefore, we propose to penalize software if it does not adhere to an open-source license (those approved in a review process by the Open Source Initiative (https://opensource.org/licenses) by allowing them only half of the total achievable points.

@@

# rigour for software implementations

>>>>

Based on four principles of a more responsible research assessment in academic hiring and promotion processes [@schonbrodt2022responsible], @gartner2022responsible suggested a concrete evaluation scheme for published manuscripts, reusable datasets, and research software.
We strongly support the increased emphasis on research software as a creditable and commendable scientific contribution.
Why are research software contributions important scientific contributions?
We would like to respond with a quote from the Science Code Manifesto [@manifesto]:
"Software is a cornerstone of science. Without software, twenty-first century science would be impossible."
and
"Software is an essential research product, and the effort to produce, maintain, adapt, and curate code must be recognized."
However, despite the heavy reliance on computational infrastructure, the current academic infrastructure does not adequately incentivise software development and, specifically, good software engineering practice [@baxter2012research].
In line with principle 3 of @schonbrodt2022responsible, we suggest that criteria for research software contributions must capture two major dimensions: rigour and impact.
Impact measures whether the scholarly effort, implementation, and dissemination actually had a visible effect on the field.
Rigour means implementing high standards and best practices for ensuring transparency, correctness, and reusability of a piece of software.
By setting a high bar of rigour in research software assessment in academic hiring and promotion, we hope to foster the creation of better software and, thus, better science.
From this perspective, we comment on some indicators of the proposed evaluation scheme for research software contributions (Table 3 of @gartner2022responsible).

@@

## ID 9: Citations/ID 5: Date of first fully functional public version

@@

Both ID 5 and 6 are difficult to ascertain because it is not always clear when updates are considered 'substantive' or software 'fully functional'.
We appreciate the importance of assessing maintenance as part of rigour.
However, the current item asks for updates that add "substantive functionality", while maintenance work (e.g., bug fixes) corresponds to patch releases.
To assess this, we propose to provide a simple checkbox, in which the author indicates whether a scientific software package is actively maintained (e.g., the software has a regular release cycle or an update within the last six months).
In addition, authors have a chance to explain why their active maintenance may be different from these guidelines.

@@

At present, whether a piece of software is open source is not evaluated in the prescreening phase.
However, an open license is central for assessing both rigour and (potential) impact, and should be part of the phase I assessment.
Above, we already discussed the different ways research software can have an impact --- not only by direct usage, but also by reusing software in other packages.
Open-source software makes broader impact more likely and is a prerequisite for full transparency and reproducibility [@peikert2021reproducible].
In addition, many aspects of rigour are impossible to evaluate for closed-source software --- for example, whether it is well-tested or bugs have been fixed.
Therefore, we propose to penalize software if it does not adhere to an open-source license (those approved in a review process by the Open Source Initiative (https://opensource.org/licenses)) by allowing them only half of the total achievable points.

@@

# Rigour for Research Software
